Automatic Matching of Bullet Lands

### 2015-12-23 imaging-paper.Rnw
done for now

\subsection{Horizontal Alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned (by horizontal shift) signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\subsection{Varying cross sections}
The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching. An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. \hh{XXX shift cross section discussion into the supplement XXX}
This situation is assessed and discussed in more detail below:

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between cross sections (c).}
\end{figure}
%\afterpage{\clearpage}

@@

Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
%\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}

>>>>

\subsection{Horizontal alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned (by horizontal shift) signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\subsection{Impact of cross sections}
The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching. An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. \hh{XXX shift cross section discussion into the supplement XXX}
This situation is assessed and discussed in more detail below:

@@

\hh{XXX Why do we use the lowest possible cross section? XXX}

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between cross sections (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
%\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
